---
title: "Schdule for iCOTS 2014 Workshop"
author: "R Pruim, D Kaplan, N Horton"
date: "July 12, 2014"
output: html_document
---



iCOTS 2014 Workshop
-------------------

## Teaching Statistics with R and RStudio

### Access to the RStudio Server at Calvin College

Each workshop participant has been given an account on an [RStudio server at Calvin College](http://acc.calvin.edu:8787).  At the workshop you will be given login credentials.

#### Working on your own machine
You are welcome to use a local installation of R and RStudio, but we will not have time during the workshop to help you install and configure the software on your machine. *We recommend that you use the RStudio server account,  even if you are familiar with R and RStudio and have access to it locally or on another server.*

If you choose to work on your own machine, be sure to install (a recent version of) the `mosaic` package using
```{r eval=FALSE}
install.packages("mosaic")    # this should be at least version 0.9
```
To get an even more recent version, use
```{r eval=FALSE}
install.packages("devtools")
require(devtools)
install_github("mosaic","rpruim")
```

### Feedback

Use this  [google doc](https://docs.google.com/document/d/1Q5a2lk9wNgUex5H60RXFgg4GjdQ9KGuJfpmnqPZHeXg/edit?usp=sharing) to provide some feed back as the workshop goes along.

### Saturday Afternoon, July 12, 2014

1. Welcome
#. Getting onto your [RStudio server](http://acc.calvin.edu:8787) account
#. An introduction to R in RStudio [Nick]
#. [Less Volume, More Creativity](Slides/LessVolume.html) -- a quick start guide to R [Randy]
#. Using RMarkdown in the classroom [Nick]
#. What are the biggest obstacles to teaching with R?   Use this [google doc](https://docs.google.com/document/d/1Q5a2lk9wNgUex5H60RXFgg4GjdQ9KGuJfpmnqPZHeXg/edit?usp=sharing) to contribute
your ideas.

### Sunday Afternoon, July 13, 2014

1. Introducing RMarkdown into your workflow [Nick and Randy]
  * [Prezi](http://prezi.com/e7wtfnacj9ub/)
  * [paper by Baumer, et al](http://escholarship.org/uc/item/90b2f5xh)
  * RMarkdown walk through
#. [Sampling and Resampling in R](Slides/SamplingResampling.html) [Danny]
#. Using your own data 

